---
title: "Final Examination"
author: "Binish Kurian Chandy"
date: "12/11/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(corrplot)
library(RColorBrewer)
library(matrixcalc)
library(moments)
library(MASS)
```

## Problem 1

```{r}
data <- read.csv("/Users/bchand005c/CUNY/DATA-605/assignment/final/data.csv", header = TRUE)
X <- data$X1
Y <- data$Y1
(x <- quantile(X, 0.75))
(y <- quantile(Y, 0.25))
(n <- nrow(data))
(ny <- nrow(subset(data, Y > y)))
```

a.   $P(X>x | Y>y)$, probability of X is higer then 3rd quantile (more then 15.825) given that Y is higher then 1st quantile (higer then 18.55)
```{r p1, echo=TRUE}
(pa <- nrow(subset(data, X > x & Y > y))/ny)
```

b.   $P(X>x, Y>y)$, probability of both X is higer then 3rd quantile (more then 15.825) and Y is higher then 1st quantile (higer then 18.55)
```{r p2, echo=TRUE}
(pb<-nrow(subset(data, X > x & Y > y))/n)
```

c.   $P(X<x | Y>y)$, probability of X is lower then 3rd quantile (less then 15.825) given that Y is higher then 1st quantile (higer then 18.55)
```{r p3, echo=TRUE}
(pc <- nrow(subset(data, X < x & Y > y))/ny)
```

### Table
```{r table, echo=TRUE}
t <- c(nrow(data[X<=x & Y<=y,]),
       nrow(data[X<=x & Y>y,]))
t <- rbind(t, c(nrow(data[X>x & Y<=y,]),
       nrow(data[X>x & Y>y,])))
t <- cbind(t, t[,1] + t[,2])
t <- rbind(t, t[1,] + t[2,])
colnames(t) <- c("<=3d quartile", ">3d quartile", "Total")
rownames(t) <- c("<=1st quartile", ">1 st quartile", "Total")
knitr::kable(t)
```

### Checking for independence
```{r independence, echo=TRUE}
X <- data$X1
Y <- data$Y1
(x <- quantile(X, 0.25))
(y <- quantile(Y, 0.25))
A <- X > x
B <- Y > y
# Calculate P(AB)
(P_AB = sum(A[B]) / length(Y))

# Calculate P(A) * P(B)
P_A = sum(A)/length(Y)
P_B = sum(B)/length(Y)
(P_A * P_B)
```
The above shows that P(AB)≠P(A)P(B), i.e. that A and B are not independent.

### Chi-Square Test
```{r chi, echo=TRUE}
(ptest <- chisq.test(table(X > x, Y > x)))
```
Since p-value is less than significane level of alpha = 0.05, the variables are statisically dependent.

## Problem 2
```{r}
train <- read.csv("/Users/bchand005c/CUNY/DATA-605/assignment/final/train.csv", header = TRUE)
summary(train)

X <- train$TotalBsmtSF
Y <- train$SalePrice

summary(X)
hist(X, xlab="TotalBsmtSF", main = "Histogram of TotalBsmtSF")
boxplot(X)

summary(Y)
hist(Y, xlab="SalePrice", main = "Histogram of Frequencies")
boxplot(Y)

plot(X,Y, xlab='TotalBsmtSF', ylab='SalePrice', main='TotalBsmtSF vs SalePrice')

Z <- train$LotFrontage 
plot(X,Z, xlab='LotFrontage', ylab='SalePrice', main='LotFrontage vs SalePrice')
```

### Correlation Matrix
```{r, warning =FALSE, echo=FALSE}
vars <- c("TotalBsmtSF", "GrLivArea", "SalePrice")
values <- train[vars]

corr_matrix <- cor(values)
round(corr_matrix, 2)

corrplot(corr_matrix, type="upper", order="hclust",  col=brewer.pal(n=8, name="RdBu"))
```
The above plot shows the correlation coefficients in colors (blue for positive correlation and red for negative
correlation).

### Test hypotheses at 80% confidence interval
### TotalBsmtSF vs GrLivArea
```{r}
cor.test(train$TotalBsmtSF, train$GrLivArea, conf.level = 0.80, method = "pearson")
```
### GrLivArea vs SalePrice
```{r}
cor.test(train$GrLivArea, train$SalePrice, conf.level = 0.80, method = "pearson")
```
### TotalBsmtSF vs SalePrice
```{r}
cor.test(train$TotalBsmtSF, train$SalePrice, conf.level = 0.80, method = "pearson")
```
Conclusion : In all the above three tests, null hypothesis is rejected in favor of alternate hypothesis.

### Would you be worried about familywise error? Why or why not?
Familywise errors are based on Type 1 or Type 2 errors. We should be worried when we observe extremely low
p value and moderate values of correlation coefficients. Here it is not the case. So I am not worried about it.

## Linear Algebra and Correlation
Correlation matrix
```{r}
corr_matrix
```
Invert the correlation matrix
```{r, echo= FALSE}
precision_matrix <- solve(corr_matrix)
precision_matrix
```
Multiply the correlation matrix by the precision matrix, and then multiply the precision matrix by the correlation matrix. 
```{r}
round(corr_matrix %*%  precision_matrix)
round(precision_matrix %*%  corr_matrix)
```
In both cases Identity matrix is obtained.

### LU Decomposition
```{r}
(LU <- lu.decomposition(precision_matrix))
```

## Calculus-Based Probability & Statistics
```{r}
X <- train$TotalBsmtSF
hist(X, main = "TotalBsmtSF")
skewness(X)
summary(X)
```
TotalBsmtSF is skewed to right.

Shifting above 0 since min value is 0
```{r}
X <- X + 1
lamda <- fitdistr(X, 'exponential')
lamda$estimate
```
The optimum of lambda is 0.0009447961

### Take 1000 samples
```{r, }
sample <- rexp(1000, lamda$estimate)
```

### Plot a histogram and compare it with a histogram of your original variable.   
```{r}
# Compare the two histograms side-by-side.
par(mfrow=c(1, 2))
hist(train$TotalBsmtSF, main="X (TotalBsmtSF)")
hist(sample, main="Exponential Distr for X")
```

### 
```{r}
# For X
quantile(train$TotalBsmtSF, probs=c(0.05, 0.95))
# For sample
quantile(sample, probs=c(0.05, 0.95))

# Generate 95% confidence interval for the emperical data:

sd = sd(X)
mean = mean(X)
n = length(X)

err = qnorm(0.975)*sd/sqrt(n)
left = mean - err
right = mean + err

cat("A 95% confidence interval for TotalBsmtSF is [", left, ",", right, "]") 
```
Conclusion : The data follow normal distribution than exponential distribution. Normal distribution would be
helpful to explain the data better than exponential distribution.

## Modeling
```{r}
train <- read.csv("/Users/bchand005c/CUNY/DATA-605/assignment/final/train.csv", header = TRUE)
df.train <- as.data.frame(train)
summary(df.train)
```
Based on summary stats, removing fields that are missing data
```{r}
df.train <- df.train %>% dplyr::select(-c(Street, Alley, LandContour, Utilities, 
                           LandSlope, Condition2, MasVnrArea, Heating, 
                           BsmtFinSF2, X2ndFlrSF, LowQualFinSF, BsmtFullBath, 
                           BsmtHalfBath, HalfBath, PoolQC, PoolArea, MiscVal, 
                           MiscFeature, Fence, ScreenPorch, Fireplaces,
                           EnclosedPorch, MoSold, YrSold))
```
Convert categorical variables to numerical values
```{r}
df.train <- df.train %>% mutate_if(is.factor, as.numeric)
```
Replace NA with 0
```{r}
df.train <- df.train %>% replace(is.na(.), 0)
```
Convert sales price to log values as feature engineering
```{r}
df.train$SalePrice <- log(df.train$SalePrice)
```
Build initial model with all selected features
```{r}
sale_price_lm <- lm(SalePrice ~ . , data=df.train)
summary(sale_price_lm)
```
Performs stepwise model selection by AIC.
```{r}
aic_lm <- stepAIC(sale_price_lm, trace=FALSE)
summary(aic_lm)
```
Check R squared value
```{r}
summary(aic_lm)$r.squared
```
High value of R squared makes our model good fit of data.

Checking the validity of model
```{r}
qqnorm(aic_lm$residuals); qqline(aic_lm$residuals)
plot(aic_lm$fitted.values, aic_lm$residuals, 
     xlab="Values", ylab="Residuals")
abline(h=0)
```
Residuals don't show any patterns so the model is valid.

Prediction
```{r}
test <- read.csv("/Users/bchand005c/CUNY/DATA-605/assignment/final/test.csv", header = TRUE)
df.test <- as.data.frame(test)
df.test <- df.test %>% dplyr::select(-c(Street, Alley, LandContour, Utilities, 
                           LandSlope, Condition2, MasVnrArea, Heating, 
                           BsmtFinSF2, X2ndFlrSF, LowQualFinSF, BsmtFullBath, 
                           BsmtHalfBath, HalfBath, PoolQC, PoolArea, MiscVal, 
                           MiscFeature, Fence, ScreenPorch, Fireplaces,
                           EnclosedPorch, MoSold, YrSold))
df.test <- df.test %>% mutate_if(is.factor, as.numeric)
df.test <- df.test %>% replace(is.na(.), 0)

pred_saleprice <- predict(aic_lm, df.test)
kaggle <- data.frame(Id=test$Id, SalePrice=pred_saleprice)
write.csv(kaggle, file = "submission.csv", row.names=FALSE)
```
Kaggle Submission
```
username : astilavista    score : 9.45492
```