---
title: "Assignment 3"
author: "Binish Kurian Chandy"
date: "9/16/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Problem Set 1
(1) What is the rank of the matrix A?   
A = 
```{r, echo=FALSE}
(A = matrix(c(1, 2, 3, 4, -1, 0, 1, 3,0, 1, -2, 1, 5, 4, -2, -3), nrow = 4, byrow = T))
```
```{r}
(rank = qr(A)$rank)
```

(2) Given an mxn matrix where m > n, what can be the maximum rank? The minimum
rank, assuming that the matrix is non-zero?
```
In a mxn matris where m > n the maximum rank will be the the minimum between m and n. 
So the maximum rank is n.

Assuming the matrix is non-zero, the minimum rank is 1.
```

(3) What is the rank of matrix B?  
B = 
```{r, echo=FALSE}
(B = matrix(c(1, 2, 1, 3, 6, 3, 2, 4, 2), nrow = 3, byrow = T))
```
```{r}
(rank = qr(B)$rank)
```

### Problem set 2

Compute the eigenvalues and eigenvectors of the matrix A. You'll need to show your
work. You'll need to write out the characteristic polynomial and show your solution.
A = 
```{r}
(A = matrix(c(1, 2, 3, 0, 4, 5, 0, 0, 6), nrow = 3, byrow = T))
```

```
Eigen values
```
$$
\lambda I_{3} =
\begin{bmatrix}
\lambda & 0 & 0 \\
0 & \lambda & 0 \\
0 & 0 & \lambda
\end{bmatrix}
$$

$$
\lambda I_{3} - A =
\begin{bmatrix}
\lambda - 1 & -2 & -3 \\
0 & \lambda - 4 & -5 \\
0 & 0 & \lambda - 6
\end{bmatrix}
$$

```
(λ−1)[(λ−4)(λ−6)−0]−2(0)+3(0)=0 

(λ−1)(λ^2−10λ+24)=0

λ^3−11λ^2+34λ−24=0 => Characteristic polynomial

(λ-1)(λ-4)(λ-6)=0

λ = 1,4,6
```

```
Eigen Vectors
```

For $\lambda_{1} = 1$

$$
\begin{bmatrix}
\lambda - 1 & -2 & -3 \\
0 & \lambda - 4 & -5 \\
0 & 0 & \lambda - 6
\end{bmatrix}
\vec{v} = \vec{0}
$$

$$
\begin{bmatrix}
0 & -2 & -3 \\
0 & -3 & -5 \\
0 & 0 & -5
\end{bmatrix}
\begin{bmatrix}
v_{1} \\
v_{2} \\
v_{3}
\end{bmatrix}=
\begin{bmatrix}
0 \\
0 \\
0
\end{bmatrix}
$$

```{r echo=TRUE}
A1 <- matrix(c(0, -2,-3, 0, -3, -5, 0, 0, -5), nrow = 3, byrow = T)
#Row reduction 
(A1[1,] <- A1[1,] / -2)

(A1[2,] <- A1[1,] *3 + A1[2,])

(A1[2,] <- A1[2,] / -.5)

(A1[1,] <- A1[1,] - A1[2,] * 1.5)

(A1[3,] <- A1[3,] + A1[2,] * 5)
A1
```

$$
(\lambda I_{3} - A)v = 
\begin{bmatrix}
0 & 1 & 0 \\
0 & 0 & 1  \\
0 & 0 & 0 
\end{bmatrix}
\begin{bmatrix}
v_{1} \\
v_{2} \\
v_{3}
\end{bmatrix}
$$

$v_{2} = 0$

$v_{3} = 0$

Let, $v_{1} = t$   
t be any real number

$$
E_{\lambda =1}= t
\begin{bmatrix}
1 \\
0 \\
0
\end{bmatrix}
$$

For $\lambda_{2} = 4$

$$
\begin{bmatrix}
3 & -2 & -3 \\
0 & 0 & -5 \\
0 & 0 & -2
\end{bmatrix}
\begin{bmatrix}
v_{1} \\
v_{2} \\
v_{3}
\end{bmatrix}=
\begin{bmatrix}
0 \\
0 \\
0
\end{bmatrix}
$$

```{r, echo=TRUE}
A2 <- matrix(c(3, -2, -3, 0, 0, -5, 0, 0, -2), nrow = 3, byrow = T)
#Row reduction
(A2[1,] <- A2[1,]/3)

(A2[2,] <- A2[2,] / -5)

(A2[1,] <- A2[1,] + A2[2,]) 

(A2[3,] <- A2[2,] * 2 + A2[3,])
A2
```

$$
(\lambda I_{3} - A)v = 
\begin{bmatrix}
1 & -0.6666667 & 0 \\
0 & 0 & 1 \\
0 & 0 & 0 
\end{bmatrix}
\begin{bmatrix}
v_{1} \\
v_{2} \\
v_{3}
\end{bmatrix}
$$

$v_{1} - 0.6666667v_{2} = 0$

$v_{3} = 0$

Let, $v_{2} =t$

$ie. 0.6666667t = v_{1}$

$$
E_{\lambda =4}= t
\begin{bmatrix}
0.6666667 \\
1 \\
0
\end{bmatrix}
$$

For $\lambda_{3} = 6$

$$
\begin{bmatrix}
5 & -2 & -3 \\
0 & 2 & -5 \\
0 & 0 & 0
\end{bmatrix}
\begin{bmatrix}
v_{1} \\
v_{2} \\
v_{3}
\end{bmatrix}=
\begin{bmatrix}
0 \\
0 \\
0
\end{bmatrix}
$$

```{r, echo=TRUE}
A3 <- matrix(c(5, -2, -3, 0, 2, -5, 0, 0, 0), nrow = 3, byrow = T)
#Row reduction
(A3[1,] <- A3[1,] / 5)

(A3[2,] <- A3[2,] / 2)

(A3[1,] <- A3[2,] * .4 + A3[1,])
A3
```

$$
(\lambda I_{3} - A)v = 
\begin{bmatrix}
1 & -0.4 & -1.6 \\
0 & 1 & -2.5 \\
0 & 0 & 0 
\end{bmatrix}
\begin{bmatrix}
v_{1} \\
v_{2} \\
v_{3}
\end{bmatrix}
$$

$v_{1} - 1.6v_{3} = 0$

$v_{2} - 2.5v_{3}= 0$

Let, $v_{3} =t$

$ie. 1.6t = v_{1}$

$2.5t = v_{2}$

$$
E_{\lambda =6}= t
\begin{bmatrix}
1.6 \\
2.5 \\
0
\end{bmatrix}
$$

check using R

```{r, echo=TRUE}
A <- matrix(c(1, 2, 3, 0, 4, 5, 0, 0, 6), nrow = 3, byrow = T)
eigen(A)
```