---
title: "Assignment-02"
author: "Binish Kurian Chandy"
date: "9/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Problem Set 1  
(1) Show that $A^TA\neq AA^T$ in general. (Proof and demonstration.)

**Proof**

Let A = $(m \times n)$ =
$$
{\begin{bmatrix}
 x_{1} \; x_{2} \; x_{3} \\ 
 y_{1} \; y_{2} \; y_{3} \\ 
 z_{1} \; z_{2} \; z_{3} 
\end{bmatrix}}
$$
$A^T = (n \times m)$ = 
$$
{\begin{bmatrix}
 x_{1} \; y_{1} \; z_{1} \\ 
 x_{2} \; y_{2} \; z_{2} \\ 
 x_{3} \; y_{3} \; z_{3} 
\end{bmatrix}}
$$
$A^TA$ =

$$
{\begin{bmatrix}
 {x_{1}}^{2} + {y_{1}}^{2} + {z_{1}}^{2} \;\;\;\;\;\;\;\;\;\;\;\;\; x_{1} x_{2} + y_{1} y_{2} + z_{1} z_{2} \;\;\;\;\;\;\; x_{1} x_{3} + y_{1} y_{3} + z_{1} z_{3} \\ 
 x_{2} x_{1} + y_{2} y_{1} + z_{2} z_{1} \;\;\;\;\;\;\; {x_{2}}^{2} + {y_{2}}^{2} + {z_{2}}^{2} \;\;\;\;\;\;\;\;\;\;\;\; x_{2} x_{3} + y_{2} y_{3} + z_{2} z_{3} \\ 
 x_{3} x_{1} + y_{3} y_{1} + z_{3} z_{1} \;\;\;\;\;\;\;\; x_{3} x_{2} + y_{3} y_{2} + z_{3} z_{2} \;\;\;\;\;\;\;\;\;\;\;\; {x_{3}}^{2} + {y_{3}}^{2} + {z_{3}}^{2} \\ 
\end{bmatrix}}
$$

$AA^T$ =
$$
{\begin{bmatrix}
 {x_{1}}^{2} + {x_{2}}^{2} + {x_{3}}^{2} \;\;\;\;\;\;\;\;\;\;\;\; x_{1} y_{1} + x_{2} y_{2} + x_{3} y_{3} \;\;\;\;\;\;\; x_{1} z_{1} + x_{2} z_{2} + x_{3} z_{3} \\ 
 y_{1} x_{1} + y_{2} x_{2} + y_{3} x_{3} \;\;\;\;\;\;\; {y_{1}}^{2} + {y_{2}}^{2} + {y_{3}}^{2} \;\;\;\;\;\;\;\;\;\;\;\;\;\; y_{1} z_{1} + y_{2} z_{2} + y_{3} z_{3} \\ 
 z_{1} x_{1} + z_{2} x_{2} + z_{3} x_{3} \;\;\;\;\;\;\;\; z_{1} y_{1} + z_{2} y_{2} + z_{3} y_{3} \;\;\;\;\;\;\;\;\;\;\;\;\;\; {z_{1}}^{2} + {z_{2}}^{2} + {z_{3}}^{2} \\ 
\end{bmatrix}}
$$
This shows that $A^TA\neq AA^T$  

**Demonstration**:

```{r}
(A <- matrix(c(1:9), nrow=3))
```

```{r}
(AT <- t(A))
```


```{r}
(ATA <- t(A)%*%A)
```

```{r}
(AAT <- A%*%t(A))
```

$A^TA\neq AA^T$ in general.  

(2) For a special type of square matrix A, we get $A^TA= AA^T$. Under what conditions could this be true? (Hint: The Identity matrix I is an example of such a matrix).

This can happen when lower and upper triangular elements without diagonal are mirror image.
Eg:

```{r}
(A <- matrix(c(5,2,4,2,0,2,4,2,5), nrow = 3, byrow = T))
```

```{r}
(AT <- t(A))
```


```{r}
(ATA <- t(A)%*%A)
```

```{r}
(AAT <- A%*%t(A))
```

```{r}
ATA == AAT
```
Since both are same, $A^TA= AA^T$

### Problem set 2

Matrix factorization is a very important problem. There are supercomputers built just to do matrix factorizations. Every second you are on an airplane, matrices are being factorized. Radars that track flights use a technique called Kalman filtering. At the heart of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear systems of equations when they track your flight using radars.

Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer.

```{r}
fun.lu <- function(A){
  if(nrow(A) != ncol(A)){
    stop("Cannot proceed with non square matrix")
  }
    else {
    n <- nrow(A)
    L <- diag(n)
    U <- A
      for (i in 1:(n - 1)) {
      for (j in (i + 1):n) {
        L[j, i] <- U[j, i] / U[i, i]
        U[j, ]  <- U[j, ] - L[j, i] * U[i, ]
      }
    }
        LU <- list("L" = L, "U" = U)
    }
}

```   

```{r}
(A <- matrix(c(1:9), nrow=3))
(fun.lu(A))
```
